Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-03 ms/data_readme.Rmd
Update dryad readme
write out selected results files

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file"
    )
  )
```

@@

Date of data collection: 2012--2014

@@

- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of ferns in Japan
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid cells (sites) for ferns in Japan
- japan_ferns_traits_lucid.csv: Morphological trait data of ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Analysis results that may be useful in future studies. Files include:
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of ferns in Japan
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of ferns in Japan
  - japan_ferns_traits.csv: Trait matrix of ferns in Japan used for functional biodiversity analysis
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of ferns in Japan used for biodiversity analysis
  - japan_ferns_biodiv.gpkg: Biodiversity statistics of ferns in Japan

@@

A list of native, non-hybrid fern specimens housed at the herbarium of the
Museum of Science and Nature, Japan was converted to a community data matrix at
four grain sizes (grid cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
per side). The original list with latitude and longitude of each specimen is not
included here to protect rare and endangered species. The 0.2 degree grain size
was selected for further analysis based on redundancy (ratio of number of
specimens to number of taxa per cell).

All taxon names are based on the Green List (http://www.rdplants.org/gl/;
English version available at
https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32)

@@

Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE
[@Nguyen2015] by combining plastid *rbcL* sequences of each taxon with a
globally sampled data matrix (Nitta et al, in prep). Next, dating analysis was
carried out using treePL v1.0 [@Smith2012] with `r n_fossil_points` fossil
calibration points after @Testo2016a. The dated phylogeny was then trimmed to
include Japanese taxa only.

@@

Data files were generated from raw data (not included here) using scripts
available at https://github.com/joelnitta/japan_ferns_biogeo.

@@

@@

`r fs::path_file(japan_ferns_comm_full_file)`: Community matrix (species x sites
matrix) of ferns in Japan. Sites correspond to 0.2 degree x 0.2 degree grid
cells covering Japan. For location of grid cells, see
japan_ferns_shape_full.gpkg. Full dataset including all grid cells (none
excluded due to low redundancy).

@@

- grids: Name of grid cell.
- Other columns: Each column is named for a Japanese fern taxon. 
  Values indicate number of specimens that were observed in each grid cell for that taxon.

@@

```{r japan_ferns_inext_results}
inext_res <- read_csv(here::here(inext_res_file))
# Check that no data are missing
assert(inext_res, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(inext_res_file)`: Output from running iNEXT [@Hsieh2016] on
specimen data for ferns of Japan. iNEXT estimates diversity (here, species
richness) and sampling completeness based on interpolation and extrapolation
from the original data.

@@

```{r results = "asis"}
  inext_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "m", "Sample size (number of individuals)",
      "method", "Method used to estimate diveristy",
      "order", "Diversity order of Hill number (here 0, equivalent to species richness)",
      "qD", "Estimated diversity",
      "qD.LCL", "Lower 95% confidence limit on estimated diversity",
      "qD.UCL", "Upper 95% confidence limit on estimated diversity",
      "SC", "Estimated sampling completeness",
      "SC.LCL", "Lower 95% confidence limit on sampling completeness",
      "SC.UCL", "Upper 95% confidence limit on sampling completeness")
  )
```

@@

```{r occ_point_data_summary}
occ_point_data_summary <- read_csv(here::here(occ_point_data_summary_file))
# Check that no data are missing
assert(occ_point_data_summary, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(occ_point_data_summary_file)`: Summary of raw specimen data for
native, non-hybrid ferns of Japan. Raw specimen data consists of one row per
specimen, with species name, latitude, longitude, and collection date (raw
specimen data not included here to protect endangered species). Raw data were
filtered to remove duplicate specimens and specimens that occurred outside of
the second-degree mesh (approximately 10 km sq grid cell) map of Japan
(http://www.biodic.go.jp/trialSystem/EN/shpddl.html)

@@

```{r results = "asis"}
  occ_point_data_summary %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens; 'occ_point_data_ferns' indicates data only includes specimens after filtering",
      "variable", "Variable. 'n_taxa' indicates number of taxa, 'n_specimens' indicates number of specimens",
      "value", "Value corresponding to each variable")
  )
```

@@

```{r redundancy_by_res}
redundancy_by_res <- read_csv(here::here(redundancy_by_res_file))
# Check that no data are missing
assert(redundancy_by_res, not_na, everything(), success_fun = success_logical)
```

@@

```{r results = "asis"}
  redundancy_by_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "res", "Degree of resolution (grid-cell side length), from 0.1 to 0.4 degrees",
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)"
  ))
```

@@

```{r japan_ferns_shape_full}
japan_ferns_shape_full <- st_read(here::here(japan_ferns_shape_full_file))
# Check that no data are missing
assert(japan_ferns_shape_full, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(japan_ferns_shape_full_file)`: Spatial data of native,
non-hybrid ferns in Japan in GeoPackage format. Compiled from raw specimen data
by binning specimens into 0.2 degree grid cells across Japan.

@@

@@

```{r results = "asis"}
  japan_ferns_shape_full %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

@@

2021-08-31

>>>>

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_dryad_file",
    "comm_ferns_dryad_file",
    "shape_ferns_dryad_file",
    "fern_traits_dryad_file",
    "japan_fern_phylogram_dryad_file",
    "japan_fern_tree_dryad_file"
    )
  )
```

@@

@@

- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_deer_range.gpkg: Distribution maps of Japanese deer (Cervus nippon) in Japan
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of native, non-hybrid ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for native, non-hybrid ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_traits_lucid.csv: Morphological trait data of native, non-hybrid ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a labeled map of Japan
- japan_map.gpkg: Map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Selected analysis results. Files include:
  - japan_ferns_biodiv.csv: Biodiversity statistics of native, non-hybrid ferns in Japan
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_shape.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan

@@

A list of native, non-hybrid fern specimens housed at the herbarium of the
Museum of Science and Nature, Japan was converted to a community data matrix at
four grain sizes (grid-cells spanning Japan, each 0.1, 0.2, 0.3, or 0.4 degrees
per side). The original list with latitude and longitude of each specimen is not
included here to protect rare and endangered species. The 0.2 degree grain size
was selected for further analysis based on redundancy (ratio of number of
specimens to number of taxa per cell).

All taxon names are based on the Green List (http://www.rdplants.org/gl/;
English version available at
https://datadryad.org/stash/dataset/doi:10.5061/dryad.4362p32).

@@

Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE v1.6.12
[@Nguyen2015] by combining plastid *rbcL* sequences of each taxon with a
globally sampled data matrix (Nitta et al, in prep). Next, dating analysis was
carried out using treePL v1.0 [@Smith2012] with `r n_fossil_points` fossil
calibration points after @Testo2016a. The dated phylogeny was then trimmed to
include Japanese taxa only.

@@

Data files were generated from raw data (not included here) using scripts
available at https://github.com/joelnitta/japan_ferns_biogeo, in particular
https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/process_raw_data.R.

@@

ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and
phylogenetic tree (Nitta et al., in prep.). For more information, see README
file contained in zipped archive.

MD5 checksum: `r tools::md5sum(here::here("data/ftol_data_release_v0.0.1.zip"))`

```{r deer-range}
deer_range <- sf::st_read(here::here(deer_range_file))
# Check that no data are missing
assert(deer_range, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(deer_range_file)`: Distribution maps of Japanese deer (Cervus
nippon) in Japan. Raw data were downloaded from the Japan Ministry of the
Environment
(https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
The data include range maps based on three types of data: range of Japanese deer
surveyed in 1978, range surveyed in 2003, and estimated range inferred from a
model including snow cover and forest type based on the 2003 survey data. Only
estimated range with <0.05 movement cost was included.

Number of variables: `r ncol(deer_range)`

Number of cases/rows: `r nrow(deer_range)`

Coordinate reference system: JGD2000

```{r results = "asis"}
  deer_range %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "range", "Source of range data. '1978' = 1978 survey, '2003' = 2003 survey, 'estimated' = range estimated from model based on 2003 survey data",
      "geom", "Vector describing shape and position of range polygon"
  ))
```

MD5 checksum: `r tools::md5sum(here::here(deer_range_file))`

```{r ja-climate-data}
ja_climate_data <- sf::st_read(here::here(ja_climate_data_file))
# Check that no data are missing
assert(ja_climate_data, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(ja_climate_data_file)`: Climate data in Japan downloaded from
WorldClim database (https://worldclim.org/data/v1.4/formats.html). Raw data were
downloaded at the 2.5 minute scale. The intersection of the raw climate data
with 0.2 degree grid-cells (japan_ferns_shape_full.gpkg) were identified,
then average value for each climatic variable was calculated for each grid-cell.

Number of variables: `r ncol(ja_climate_data)`

Number of cases/rows: `r nrow(ja_climate_data)`

Coordinate reference system: JGD2000

```{r ja-climate-data-cols, results = "asis"}
  ja_climate_data %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "temp", "Mean annual temperature, in units of 10 * degrees celsius (BIO1)",
      "temp_season", "Temperature seasonality (standard deviation ×100) (BIO4)",
      "precip", "Annual precipitation (mm) (BIO12)",
      "precip_season", "Precipitation seasonlaity (coefficient of variation) (BIO15)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

MD5 checksum: `r tools::md5sum(here::here(ja_climate_data_file))`

@@

`r fs::path_file(japan_ferns_comm_full_file)`: Community matrix (species x sites
matrix) of native, non-hybrid ferns in Japan. Sites correspond to 0.2 degree x
0.2 degree grid-cells covering Japan. For location of grid-cells, see
japan_ferns_shape_full.gpkg. Full dataset including all grid-cells (none
excluded due to low redundancy).

@@

- grids: Name of grid-cell.
- Other columns: Each column is named for a Japanese fern taxon. 
  Values indicate number of specimens that were observed in each grid-cell for that taxon.

@@

```{r japan-ferns-inext-results}
inext_res <- read_csv(here::here(inext_res_file))
# Check that no data are missing
assert(inext_res, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(inext_res_file)`: Output from running iNEXT [@Hsieh2016] on
specimen data of native, non-hybrid ferns in Japan. iNEXT estimates diversity
(here, species richness) and sampling completeness based on interpolation and
extrapolation from the original data.

@@

```{r japan-ferns-inext-results-cols, results = "asis"}
  inext_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "m", "Sample size (number of individuals)",
      "method", "Method used to estimate diveristy",
      "order", "Diversity order of Hill number (here 0, equivalent to species richness)",
      "qD", "Estimated diversity",
      "qD.LCL", "Lower 95% confidence limit on estimated diversity",
      "qD.UCL", "Upper 95% confidence limit on estimated diversity",
      "SC", "Estimated sampling completeness",
      "SC.LCL", "Lower 95% confidence limit on sampling completeness",
      "SC.UCL", "Upper 95% confidence limit on sampling completeness")
  )
```

@@

```{r occ-point-data-summary}
occ_point_data_summary <- read_csv(here::here(occ_point_data_summary_file))
# Check that no data are missing
assert(occ_point_data_summary, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(occ_point_data_summary_file)`: Summary of raw specimen data for
native, non-hybrid ferns of Japan. Raw specimen data consists of one row per
specimen, with species name, latitude, longitude, and collection date (raw
specimen data not included here to protect endangered species). Raw data were
filtered to remove duplicate specimens and specimens that occurred outside of
the second-degree mesh (approximately 10 km sq grid-cell) map of Japan
(http://www.biodic.go.jp/trialSystem/EN/shpddl.html)

@@

```{r occ-point-data-summary-cols, results = "asis"}
  occ_point_data_summary %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "dataset", "Dataset name. 'occ_point_data_ferns_unfiltered' indicates data includes all specimens; 'occ_point_data_ferns' indicates data only includes specimens after filtering",
      "variable", "Variable. 'n_taxa' indicates number of taxa, 'n_specimens' indicates number of specimens",
      "value", "Value corresponding to each variable")
  )
```

@@

```{r redundancy-by-res}
redundancy_by_res <- read_csv(here::here(redundancy_by_res_file))
# Check that no data are missing
assert(redundancy_by_res, not_na, everything(), success_fun = success_logical)
```

@@

```{r redundancy-by-res-cols, results = "asis"}
  redundancy_by_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "res", "Degree of resolution (grid-cell side length), from 0.1 to 0.4 degrees",
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)"
  ))
```

@@

```{r japan-ferns-shape-full}
japan_ferns_shape_full <- st_read(here::here(japan_ferns_shape_full_file))
# Check that no data are missing
assert(japan_ferns_shape_full, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(japan_ferns_shape_full_file)`: Location of grid-cells (sites)
for native, non-hybrid ferns in Japan in GeoPackage format. Compiled from raw
specimen data by binning specimens into 0.2 degree grid-cells across Japan. Full
dataset including all grid-cells (none excluded due to low redundancy).

@@

Coordinate reference system: JGD2000

@@

```{r japan-ferns-shape-full-cols, results = "asis"}
  japan_ferns_shape_full %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

```{r japan-ferns-traits-lucid}
japan_ferns_traits_lucid <- read_csv(here::here(raw_trait_data_file))
# Check that no data are missing
assert(japan_ferns_traits_lucid, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(raw_trait_data_file)`: Morphological trait data of native,
non-hybrid ferns in Japan, originally formatted for Lucid dichotomous key
software (https://www.lucidcentral.org/). For more information on codes used for
scoring of traits, see https://help.lucidcentral.org/lucid/scoring-the-key/

Number of variables: `r ncol(japan_ferns_traits_lucid)`

Number of cases/rows: `r nrow(japan_ferns_traits_lucid)`

Variable list: 'taxon' indicates taxon name. Other columns correspond to traits.
If the column name includes 'number' it is numeric, otherwise categorical.

Categorical trait codes as follows:
- 0 = absent
- 1 = common
- 2 = rare
- 3 = uncertain
- 4 = common and misinterpreted
- 5 = rare and misinterpreted
- 6 = not scoped

Numeric trait data formatted as a series of numbers separated by colons (':'),
unless missing (0 or 3). For example, 1:9.4:12:14:16. The starting '1:' does not
mean anything. The next series of numbers correspond to the outside minimum,
normal minimum, normal maximum and outside maximum. For example, a plant leaf
may normally be in the range of 4-5 cm, but sometimes as low as 2 cm or as high
as 6 cm. This would be written 2:4:5:6.

Missing data codes: '0' indicates trait is missing; '3' indicates uncertain.

MD5 checksum: `r tools::md5sum(here::here(raw_trait_data_file))`

```{r japan-map-points}
japan_map_points <- read_csv(here::here(japan_map_points_file))
```

`r fs::path_file(japan_map_points_file)`: Latitude and longitude of points used
to produce a labeled map of Japan. Some points were originally generated using
script at https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/geocode.R

Number of variables: `r ncol(japan_map_points)`

Number of cases/rows: `r nrow(japan_map_points)`

```{r japan-map-points-cols, results = "asis"}
  japan_map_points %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "query", "Name of point of interest",
      "lat", "Latitude",
      "lon", "Longitude",
      "lat_end", "Ending latitude (only for features that are lines, not points)",
      "lon_end", "Ending longitude (only for features that are lines, not points)"
  ))
```

Missing data codes: No data entered between commas in CSV.

MD5 checksum: `r tools::md5sum(here::here(japan_map_points_file))`

```{r japan-map}
japan_map <- sf::st_read(here::here(japan_map_file))
# Check that no data are missing
assert(japan_map, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(japan_map_file)`: Map of land area of Japan. Raw data downloaded from
Geospatial Information Authority of Japan under the Creative Commons Attribution
License v4.0 (https://www.gsi.go.jp/kankyochiri/gm_japan_e.html). All polygons
were combined into a single polygon and CRS set to JGD2000.

Number of variables: `r ncol(japan_map)`

Number of cases/rows: `r nrow(japan_map)`

```{r japan-map-cols, results = "asis"}
  japan_map %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "geom", "Vector describing shape and position of land area polygon"
  ))
```

MD5 checksum: `r tools::md5sum(here::here(japan_map_file))`

```{r protected-areas}
protected_areas <- st_read(here::here(protected_areas_file))
# Check that no data are missing
assert(protected_areas, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(protected_areas_file)`: Spatial of protected areas in Japan
in GeoPackage format. Original spatial data downloaded from the Japan Ministry
of the Environment
(https://www.biodic.go.jp/biodiversity/activity/policy/map/map17/index.html) and
Ministry of Land, Infrastructure, Transport and Tourism
(https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-A45.html). Categorization of
protected areas generally followed @Kusumoto2017: either "high" (no human
activities allowed at all) or "medium" status (some economic activities allowed
by permit); other areas not meeting these criteria or not including plants were
not considered. Any overlapping areas within the same protection level were
combined, and any areas overlapping between "medium" and "high" were considered
only "high". Code used to categorize areas available at
https://github.com/joelnitta/japan_ferns_biogeo/blob/main/R/process_raw_data.R

Number of variables: `r ncol(protected_areas)`

Number of cases/rows: `r nrow(protected_areas)`

Coordinate reference system: JGD2000

```{r protected-areas-cols, results = "asis"}
  protected_areas %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "status", "Protection status ('high' or 'medium')",
      "geom", "Vector describing shape and position of protected area polygon"
  ))
```

MD5 checksum: `r tools::md5sum(here::here(protected_areas_file))`

```{r plastome-calibration-dates}
plastome_calibration_dates <- read_csv(here::here(calibration_dates_file))
```

`r fs::path_file(calibration_dates_file)`: Fossil calibration points used for
phylogenetic dating analysis with treePL [@Smith2012], based on @Testo2016a.
MRCA = "most recent common ancestor".

Number of variables: `r ncol(plastome_calibration_dates)`

Number of cases/rows: `r nrow(plastome_calibration_dates)`

```{r plastome-calibration-dates-cols, results = "asis"}
  plastome_calibration_dates %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "Clade", "Name of clade",
      "Stem_Crown", "Type of node. 'crown' (node corresponding to MRCA of extant taxa only) or 'stem' (node corresponding to split between MRCA of extant taxa and its sister)",
      "Fossil", "Name of fossil",
      "Age", "Age of fossil",
      "Age_type", "Type of age. 'min' for minimum age, or 'fixed' for fixed age",
      "Citation", "Citation for the fossil",
      "taxon_1", "Name of first taxon that is used to identify MRCA",
      "taxon_2", "Name of second taxon that is used to identify MRCA",
      "note", "Notes, in particular explaining any differences with Testo and Sundue (2016)"
  ))
```

Missing data codes: No data entered between commas in CSV.

MD5 checksum: `r tools::md5sum(here::here(calibration_dates_file))`

```{r japan-ferns-biodiv-dryad}
japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
# Check that no data are missing
assert(japan_ferns_biodiv_dryad, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(japan_ferns_biodiv_dryad_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. A
single outlier grid-cell with extremely low taxon richness and extremely high %
apomictic taxa was excluded. Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Data used for spatial models
in Nitta et al. 2021.

Number of variables: `r ncol(japan_ferns_biodiv_dryad)`

Number of cases/rows: `r nrow(japan_ferns_biodiv_dryad)`

```{r japan-ferns-biodiv-dryad-cols, results = "asis"}
  japan_ferns_biodiv_dryad %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "lat", "Latitude of grid-cell centroid",
      "long", "Longitude of grid-cell centroid",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "fd_obs_z", "Z-score (standard effect size) of observed functional diversity (FD)",
      "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
      "pe_obs_z", "Z-score of observed phylogenetic endemisim (PE)",
      "rfd_obs_z", "Z-score of observed relative FD",
      "rpd_obs_z", "Z-score of observed relative PD",
      "richness_obs_p_upper", "Richness percentile",
      "fd_obs_p_upper", "FD percentile",
      "pd_obs_p_upper", "PD percentile",
      "pe_obs_p_upper", "PE percentile",
      "pd_signif", "Significance of PD compared to null distribution, two-sided test at alpha 0.05",
      "rpd_signif", "Significance of RPD compared to null distribution, two-sided test at alpha 0.05",
      "fd_signif", "Significance of FD compared to null distribution, two-sided test at alpha 0.05",
      "rfd_signif", "Significance of RFD compared to null distribution, two-sided test at alpha 0.05",
      "pe_signif", "Significance of PE compared to null distribution, two-sided test at alpha 0.05",
      "taxonomic_cluster", "Membership in cluster based on taxonomic distances",
      "phylo_cluster", "Membership in cluster based on phylogenetic distances",
      "endem_type", "Endemism type determined by CANAPE",
      "lat_area", "Area of 1-degree latitudinal band around grid-cell",
      "temp", "Annual mean temperature, in units of degrees celsius * 10",
      "temp_season", "Temperature seasonality",
      "precip", "Annual precipitation, measured in mm",
      "precip_season", "Precipitation seasonality",
      "percent_apo", "Percent apomictic taxa"
    ))
```

MD5 checksum: `r tools::md5sum(here::here(japan_ferns_biodiv_dryad_file))`

```{r comm-ferns-dryad-file}
comm_ferns_dryad <- read_csv(here::here(comm_ferns_dryad_file))
# Check that no data are missing
assert(comm_ferns_dryad, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(comm_ferns_dryad_file)` (contained in "results.zip"): Community
matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for
biodiversity analysis. Same as japan_ferns_comm_full.csv, but grid-cells with
redundancy <0.1 (indicating under-sampling) excluded.

Number of variables: `r ncol(comm_ferns_dryad)`

Number of cases/rows: `r nrow(comm_ferns_dryad)`

- grids: Name of grid-cell.
- Other columns: Each column is named for a Japanese fern taxon. 
  Values indicate number of specimens that were observed in each grid-cell for that taxon.

MD5 checksum: `r tools::md5sum(here::here(comm_ferns_dryad_file))`

```{r shape-ferns-dryad-file}
shape_ferns_dryad <- sf::st_read(here::here(shape_ferns_dryad_file))
# Check that no data are missing
assert(shape_ferns_dryad, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(shape_ferns_dryad_file)` (contained in "results.zip"): Location
of grid-cells (sites) for native, non-hybrid ferns in Japan used for
biodiversity analysis. Same as japan_ferns_shape_full.gpkg, but grid-cells with
redundancy <0.1 (indicating under-sampling) excluded.

Number of variables: `r ncol(shape_ferns_dryad)`

Number of cases/rows: `r nrow(shape_ferns_dryad)`

Coordinate reference system: JGD2000

```{r shape-ferns-dryad-file-cols, results = "asis"}
  shape_ferns_dryad %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

MD5 checksum: `r tools::md5sum(here::here(shape_ferns_dryad_file))`

```{r fern-traits-dryad-file}
fern_traits_dryad <- read_csv(here::here(fern_traits_dryad_file))
```

`r fs::path_file(fern_traits_dryad_file)` (contained in "results.zip"): Trait
matrix of native, non-hybrid ferns in Japan used for functional biodiversity
analysis.

Number of variables: `r ncol(fern_traits_dryad)`

Number of cases/rows: `r nrow(fern_traits_dryad)`

Variable list: 'taxon' indicates taxon name. Other columns correspond to traits.
'frond_width', 'stipe_length', 'number_pinna_pairs' are numeric; others are
binary (0 or 1)

Missing data codes: Missing data coded as 'NA'.

MD5 checksum: `r tools::md5sum(here::here(fern_traits_dryad_file))`

```{r japan-fern-phylogram-dryad-file}
japan_fern_phylogram_dryad <- ape::read.tree(here::here(japan_fern_phylogram_dryad_file))
```

`r fs::path_file(japan_fern_phylogram_dryad_file)` (contained in
"results.zip"): Maximim-likelihood phylogeny of native, non-hybrid ferns in
Japan in newick format inferred with maximum likelihood IQ-TREE v1.6.12
[@Nguyen2015]. Values at nodes are bootstrap support values calculated using
1000 replicates of ultrafast bootstrap [@Nguyen2015].

Number of tips: `r ape::Ntip(japan_fern_phylogram_dryad)`

MD5 checksum: `r tools::md5sum(here::here(japan_fern_phylogram_dryad_file))`

```{r japan-fern-dryad-file}
japan_fern_tree_dryad <- ape::read.tree(here::here(japan_fern_tree_dryad_file))
```

`r fs::path_file(japan_fern_tree_dryad_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r
n_fossil_points` fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015].

Number of tips: `r ape::Ntip(japan_fern_tree_dryad)`

MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_dryad_file))`

@@

2021-09-03
